---
title: 'A4-task2: Parameteer Estimation'
author: "Kerry Nixon"
date: "3/5/2021"
output: 
  html_document:
    code_folding: 'hide'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)

library(tidyverse)
library(here)
library(boot)
library(gt)
library(patchwork)
library(broom)
library(nlstools)
library(janitor)
```


```{r}
# Read in data

catch <- read.csv(here("data/fish_catch.csv"), skip = 2) %>% 
  clean_names() %>% 
  drop_na() %>% 
  mutate(wild_catch = as.numeric(wild_catch)) %>% 
  mutate(year_index = as.numeric(year) - 1950)

# Exploratory Data Viz
ggplot(data = catch, aes(x = year_index, y = wild_catch)) +
  geom_point() +
  theme_minimal() +
  labs(x = "time (years, begin 1950)", y = "wild fish catch (mm tons)")

# Log tansformed version:
ggplot(data = catch, aes(x = year_index, y = log(wild_catch))) +
  geom_point() +
  theme_minimal() +
  labs(x = "time (years, begin 1950)", y = "wild fish catch (mm tons)")
```


The wild fish catch looks to increase logistically, with an upper bound limiting growth in catch appearing in data in the last 25 years of observations. 

Recall the logistic growth equation: 

$P(t)=\frac{K}{1+Ae^{-kt}}$, where

- $K$ is the upper bound on wild catch
- $A$ is $\frac{K-P_0}{P_0}$
- $k$ is the growth rate constant


```{r}
# Get only up to 40 years of catch data & log transform the catch tonnage
# We do this so we can estimate the growth rate constant (k) *during exponential growth phase)
catch_exp <- catch %>% 
  filter(year_index < 40) %>% 
  mutate(ln_wild_catch = log(wild_catch))
  
# Model linear to get *k* estimate (the slope of this linear equation is an estimate of the growth rate constant):
lm_k <- lm(ln_wild_catch ~ year_index, data = catch_exp)
lm_k

# Coefficient (k) ~ 0.036
```
Now we have initial estimate for *k* (0.036), and we can estimate *K* ~ 90 and *A* ~ 4.2 as starting points for iterative algorithms trying to converge on the parameters. 

We'll estimate the parameters using nonlinear least squares (NLS): 

```{r}
df_nls <- nls(wild_catch ~ K/(1 + A*exp(-k*year_index)),
              data = catch,
              start = list(K = 90, A = 4.2, k = 0.036),
              trace = TRUE)

# See the model summary (null hypothesis: parameter value = 0)
summary(df_nls)

# Use broom:: functions to get model outputs in tidier format: 
model_out <- broom::tidy(df_nls) # Use View(model_out) to see the tidy model output.
model_out

# Want to just get one of these?  
A_est <- model_out$estimate[2] # Gets the first rown in the estimate column.
```

Our model with estimated parameters is:
$$P(t) = \frac{100.28}{1+4.32e^{-0.07t}}$$
```{r}
# Make predictions for the population at all of those times (time) in the original df: 
catch_predict <- predict(df_nls)

# Bind predictions to original data frame:
catch_complete <- data.frame(catch, catch_predict)

# Plot them all together:
ggplot(data = catch_complete, aes(x = year_index, y = wild_catch)) +
  geom_point() +
  geom_line(aes(x = year_index, y = catch_predict)) +
  theme_minimal()

```


## Data Citation:

Global wild fish catch and aquaculture production, compiled by Earth Policy Institute with 1950-2010 from U.N. Food and Agriculture Organization (FAO), Global Capture Production and Global Aquaculture Production, electronic databases, at www.fao.org/fishery/topic/16140/en.


