---
title: "a4-task1 - agglomerative hierarchical clustering"
author: "Kerry Nixon"
date: "3/5/2021"
output: 
  html_document:
    code_folding: 'hide'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(palmerpenguins)
library(NbClust)
library(cluster)
library(factoextra)
library(dendextend)
library(ggdendro)
library(naniar)
library(stringr)
```


### Hierarchical Clustering

For Task 1, use hierarchical clustering by complete linkage to create a dendrogram showing multivariate clustering for water chemistry by site. 

To perform hierarchical clustering by site, you’ll want to make a data frame that has a single summary row per site (e.g. based on means from all observations at that site), then calculate the euclidean distance before performing complete linkage agglomerative hierarchical clustering. 


```{r}
# Reading in the data and replacing "-999" with "NA" and clean_names()
stream_chem_raw <- read.csv(here("data/sbc_lter_registered_stream_chemistry.csv")) %>% 
  replace_with_na_all(condition = ~.x == -999) %>% 
  clean_names()

# Data Wrangling:

# Drop rows where any of the four size measurements are missing
stream_chem_summ <- stream_chem_raw %>% 
  group_by(site_code) %>% 
  summarize(nh4_u_m = mean(nh4_u_m, na.rm = TRUE),
            no3_u_m = mean(no3_u_m, na.rm = TRUE),
            po4_u_m = mean(po4_u_m, na.rm = TRUE),
            tdn_u_m = mean(tdn_u_m, na.rm = TRUE),
            tdp_u_m = mean(tdp_u_m, na.rm = TRUE),
            tpc_u_m = mean(tpc_u_m, na.rm = TRUE),
            tpn_u_m = mean(tpn_u_m, na.rm = TRUE),
            tpp_u_m = mean(tpp_u_m, na.rm = TRUE),
            tss_mgper_liter = mean(tss_mgper_liter, na.rm = TRUE),
            spec_cond_u_spercm = mean(spec_cond_u_spercm, na.rm = TRUE)) %>% 
  drop_na()

# Exploratory Data Visualization

# nh4 vs po4 exploratory plot:
ggplot(stream_chem_summ) +
  geom_point(aes(x = nh4_u_m, 
                 y = po4_u_m),
             size = 3,
             alpha = 0.7)

# tdn vs tpn exploratory plot: 
ggplot(stream_chem_summ) +
  geom_point(aes(x = tdn_u_m, 
                 y = tpn_u_m),
             size = 3,
             alpha = 0.7)


# Data wrangling:

# Scale numeric variables
stream_scale <- stream_chem_summ %>% 
  select(ends_with("m")) %>% 
  scale()

# Update to add rownames (country name) from wb_ghg_20
rownames(stream_scale) <- stream_chem_summ$site_code

# Compute dissimilarity values (Euclidean distances):
euc_dist <- dist(stream_scale, method = "euclidean")

# Checking:
#View(euc_dist)

# Hierarchical clustering (complete linkage)
hc_complete <- hclust(euc_dist, method = "complete" )

# Plot it (base plot):
plot(hc_complete, cex = 0.6, hang = -1)

# Convert to class dendrogram
dend_complete <- as.dendrogram(hc_complete)

# Data Visualization - Stream chemistry dendrogram
ggdendrogram(hc_complete, 
             rotate = TRUE) +
  theme_minimal() +
  labs(x = "Stream Site Code",
       y = "Euclidian Distance")

# COOL. Then you can customize w/ usual ggplot tools. 

```







## Data Citation: 

Santa Barbara Coastal LTER and J. Melack. 2019. SBC LTER: Land: Stream chemistry in the Santa Barbara Coastal drainage area, ongoing since 2000 ver 16. Environmental Data Initiative. https://doi.org/10.6073/pasta/67a558a24ceed9a0a5bf5e46ab841174. 



